pirouette: Determining the error BEAST2 makes in inferring a phylogeny

### 2019-03-11 article.tex
Remove table with results

\iffalse
\begin{table}
\centering
  \begin{tabular}{|@{}c|p{2.8cm}|p{6cm}@{}|}
    \hline
    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
    \hline
    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
    $\mathit{s}$ & Site model & Nucleotide substitution model\\
    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor \\
    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
    \hline
    $\mathit{P}$ & Generative prior & Diversification model under which the 'true' tree is simulated \\
    $\mathit{A}$ & Alignment model & Alignment simulation model. It is a combination of $(\mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s})$, where $\mathit{c}$ = strict \\
    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference \\
    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in this article.
  }
  \label{tab:definitions}
  \richel{I've some set notations in some spots, as I am unsure if you like it}
  \giovanni{It would be nice to specify (in the first part of the table) 
    the basic elements composing all the elements of the second part of the table. 
    This could be probably done for "Alignment parameters", "Inference conditions" 
    and "Twinning parameters".
  }
  \richel{
    I feel the next table (which was also your suggestion), does so
    at the right amount of detail
  }
  \giovanni{
    How do we specify what constitutes 'twinning parameters'?
  }
  \richel{Perhaps you'd enjoy my definition?}
\end{table}
\fi

\iffalse
% previous version
\begin{table}
\centering
  \begin{tabular}{|@{}c|p{2.8cm}|@{}l|p{5.5cm}@{}|}
    \hline
    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
    \hline
    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
    $\mathit{s}$ & Site model & Nucleotide substitution model\\
    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor\\
    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
    \hline
    $\mathit{G}$ & Generative prior & Diversification model the 'true' tree is simulated with \\
    $\mathit{A}$ & Alignment model & Alignment simulation model, $\mathit{A} \ni \{ \mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s} \}$, in which $\mathit{c}$ = strict \\
    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in inference \\
    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in this article.
  }
  \label{tab:definitions}
  \richel{I've some set notations in some spots, as I am unsure if you like it}
  \giovanni{It would be nice to specify (in the first part of the table) 
    the basic elements composing all the elements of the second part of the table. 
    This could be probably done for "Alignment parameters", "Inference conditions" 
    and "Twinning parameters".
  }
  \richel{
    I feel the next table (which was also your suggestion), does so
    at the right amount of detail
  }
  \giovanni{
    How do we specify what constitutes 'twinning parameters'?
  }
  \richel{Perhaps you'd enjoy my definition?}
\end{table}
\fi

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{workflow.png}
  \caption{
    \texttt{pirouette} pipeline. 
    The pipeline starts from a true phylogeny (1a) simulated under the selected generative prior $\mathit{P}$, that is converted to an alignment (2a) according to the alignment model $\mathit{A}$. The user supplies one or more experiments. Every experiment $\mathit{X}$ is defined as a combination of an inference model $\mathit{I}$ and the conditions $\mathit{C}$ under which it is actually used in the inference. Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used in the next step (3a).
    Selected experiments use the alignment (2a) and their inference model $\mathit{I}$ to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a). The posterior trees are compared to the true phylogeny (1a) using the selected error measure $\mathit{E}$, in our case the nLTT statistic [\cite{janzen2015approximate}]. Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
    of error values (5a). Optionally, a twin phylogeny (1b) can be generated from the true phylogeny (1a), which then results in an error distribution (5b) with similar intermediate steps.
  }
  \label{fig:pipeline}
  \giovanni{add reference to the generative prior P (or G) in the figure}
  \giovanni{I think that a list of possible models should replace the "experiment" cake in the middle, then C should be drawn on the arrow connecting the figure in the middle to 3a. "I" should replace the circled x of passage 3a. An additional C can be drawn alongside the arrow 2a -> 3a. After that a rectangle can be drawn around the whole middle section (including 3a, 3b and the middle C) to denote that this is phase "X".}
  \giovanni{We might also need an arrow going from the central figure to the top arrow (the one with "T") to specify that the twin is generated according to the chosen inference model.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | r | l | l | l | l | }
\hline
\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
\cline{2-5}
                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
\hline
1  & $0.02$  & $0.0 - 0.27$  &         &               \\
2  & $0.02$  & $0.0 - 0.08$  &         &               \\
3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
\hline
\end{tabular}
\caption{
  \richel{redo}
  Results of the examples. 
  For the true and twin tree,
  the median and range (1st and 3rd quartile) of the errors 
  is shown. Experiment 4 is split in two rows
  by either the candidate ('C') or generative model ('G')
}
\label{tab:results}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

>>>>

\iffalse
\begin{table}
\centering
  \begin{tabular}{|@{}c|p{2.8cm}|p{6cm}@{}|}
    \hline
    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
    \hline
    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
    $\mathit{s}$ & Site model & Nucleotide substitution model\\
    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor \\
    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
    \hline
    $\mathit{P}$ & Generative prior & Diversification model under which the 'true' tree is simulated \\
    $\mathit{A}$ & Alignment model & Alignment simulation model. It is a combination of $(\mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s})$, where $\mathit{c}$ = strict \\
    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference \\
    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in this article.
  }
  \label{tab:definitions}
\end{table}
\fi

\iffalse
% previous version
\begin{table}
\centering
  \begin{tabular}{|@{}c|p{2.8cm}|@{}l|p{5.5cm}@{}|}
    \hline
    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
    \hline
    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
    $\mathit{s}$ & Site model & Nucleotide substitution model\\
    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor\\
    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
    \hline
    $\mathit{G}$ & Generative prior & Diversification model the 'true' tree is simulated with \\
    $\mathit{A}$ & Alignment model & Alignment simulation model, $\mathit{A} \ni \{ \mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s} \}$, in which $\mathit{c}$ = strict \\
    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in inference \\
    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in this article.
  }
  \label{tab:definitions}
\end{table}
\fi

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{workflow.png}
  \caption{
    \texttt{pirouette} pipeline. 
    The pipeline starts from a true phylogeny (1a) simulated under the selected generative prior $\mathit{P}$, that is converted to an alignment (2a) according to the alignment model $\mathit{A}$. The user supplies one or more experiments. Every experiment $\mathit{X}$ is defined as a combination of an inference model $\mathit{I}$ and the conditions $\mathit{C}$ under which it is actually used in the inference. Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used in the next step (3a).
    Selected experiments use the alignment (2a) and their inference model $\mathit{I}$ to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a). The posterior trees are compared to the true phylogeny (1a) using the selected error measure $\mathit{E}$, in our case the nLTT statistic [\cite{janzen2015approximate}]. Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
    of error values (5a). Optionally, a twin phylogeny (1b) can be generated from the true phylogeny (1a), which then results in an error distribution (5b) with similar intermediate steps.
  }
  \label{fig:pipeline}
  \giovanni{
    add reference to the generative prior P (or G) in the figure
  }
  \giovanni{
    I think that a list of possible models should replace the "experiment" 
    cake in the middle, then C should be drawn on the arrow 
    connecting the figure in the middle to 3a. 
    "I" should replace the circled x of passage 3a. 
    An additional C can be drawn alongside the arrow 2a -> 3a. 
    After that a rectangle can be drawn around the whole middle section 
    (including 3a, 3b and the middle C) to denote that this is phase "X".}
  \giovanni{We might also need an arrow going from the central figure to the top arrow (the one with "T") to specify that the twin is generated according to the chosen inference model.}
\end{figure}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\iffalse
\begin{figure}[h]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.4\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.13\textheight]{example_4_true_alignment.png}
    };   
    \node[state] (CG) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_true_posterior_gen.png}
    };   
    \node[state] (DG) [below of = CG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_error_violin_gen.png}
    };   
    \node[state] (CB) [below right of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_true_posterior_best.png}
    };   
    \node[state] (DB) [below of = CB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_error_violin_best.png}
    };   
    \node[state] (AT) [right of = A, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_tree.png}
    };   
    \node[state] (BT) [below of = AT, rectangle] {
      \includegraphics[height=0.13\textheight]{example_4_twin_alignment.png}
    };   
    \node[state] (CTG) [below of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_gen.png}
    };   
    \node[state] (DTG) [below of = CTG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_gen.png}
    };   
    \node[state] (CTB) [below right of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_best.png}
    };   
    \node[state] (DTB) [below of = CTB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_best.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (CG)
      (CG) edge [anchor = south] node {} (DG)
      (B) edge [anchor = southeast] node {} (CB)
      (CB) edge [anchor = south] node {} (DB)
      (A) edge [anchor = east] node {} (AT)
      (AT) edge [anchor = south] node {} (BT)
      (BT) edge [anchor = southeast] node {} (CTG)
      (CTG) edge [anchor = south] node {} (DTG)
      (BT) edge [anchor = south] node {} (CTB)
      (CTB) edge [anchor = south] node {} (DTB)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_4_full_pipeline}
  \caption{Example 4: full pipeline}
\end{figure}
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
